---
title: "Market Basket Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
```

## Data processing

### Load the data

Load the data downloaded from http://archive.ics.uci.edu/ml/datasets/online+retail 

```{r}
online_retail_data <- read_excel('./data/online-retail.xlsx')
online_retail_data <- online_retail_data[complete.cases(online_retail_data), ]
str(online_retail_data)
```

### Data cleaning:

We have negative quantities, we can remove them:
```{r}
online_retail_data <- filter(online_retail_data, Quantity > 0)
```

We have prices as 0, we can remove them as well:
```{r}
online_retail_data <- filter(online_retail_data, UnitPrice > 0)
```

Based on the below code we can safely say that we don't have any NA values in the dataset:
```{r}
which(is.na(online_retail_data))
```

### Convert the data types and formats of the fields:

Here the columns CustomerID, Country, StockCode can be changed to factor. 

```{r}
online_retail_data$Country <- as.factor(online_retail_data$Country)
online_retail_data$CustomerID <- as.factor(online_retail_data$CustomerID)
online_retail_data$StockCode <- as.factor(online_retail_data$StockCode)
```

Similarly we can construct the PurchaseTime and PurchaseTimeInHours of the goods from invoice date and convert the InvoiceDate to Date datatype. 
```{r}
online_retail_data$PurchaseTime <- format(online_retail_data$InvoiceDate,"%H:%M:%S")
online_retail_data$PurchaseTimeInHours <- format(online_retail_data$InvoiceDate,"%H")
online_retail_data$InvoiceDate <- as.Date(online_retail_data$InvoiceDate)
```

We can construct total cose of the oder
```{r}
online_retail_data$TotalCost <- online_retail_data$Quantity * online_retail_data$UnitPrice
```

### Save the data into csv format:
```{r}
write_csv(online_retail_data, file="./data/online-retail.csv")
```

## Exploratory Analysis:

### Top 10 countries based on orders

```{r}
bar_graph_data <- online_retail_data %>% 
  group_by(Country) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
bar_graph_data <- head(bar_graph_data, n=10)

ggplot(bar_graph_data, aes(x=count, y=Country, fill=Country)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top 10 countries based on orders") +
  xlab("Number of orders")
```

### Purchases on the hour of the day

```{r}
  ggplot(online_retail_data, aes(x=PurchaseTimeInHours)) + 
  geom_histogram(stat="count", fill="green") +
  ggtitle("Purchases on the hour of the day") +
  xlab("Time in hours") +
  ylab("Number of purchases")
```
### Top 10 customers based on number of orders placed

```{r}
bar_graph_data <- online_retail_data %>% 
  group_by(CustomerID) %>% 
  summarize(Count = n()) %>% 
  arrange(desc(Count))
bar_graph_data <- head(bar_graph_data, n=10)

ggplot(bar_graph_data, aes(x=Count, y=CustomerID, fill=CustomerID)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top 10 customers based on orders") +
  xlab("Number of orders")
```

### Top 10 items sold

```{r}
bar_graph_data <- online_retail_data %>% 
  group_by(StockCode) %>% 
  summarize(Count = n()) %>% 
  arrange(desc(Count))
bar_graph_data <- head(bar_graph_data, n=10)
bar_graph_data

ggplot(bar_graph_data, aes(x=Count, y=StockCode, fill=StockCode)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top 10 items based on orders") +
  xlab("Number of orders")
```

### Top 20 expensive orders

```{r}
bar_graph_data <- online_retail_data %>% 
  arrange(desc(TotalCost))
bar_graph_data <- head(bar_graph_data, n=20)

ggplot(bar_graph_data, aes(x=TotalCost, y=InvoiceNo, fill=InvoiceNo)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top 20 expensive orders") +
  xlab("Total Cost")
```

### Top 10 least expensive items
```{r}
bar_graph_data <- online_retail_data %>% 
  arrange(desc(TotalCost))
bar_graph_data <- tail(bar_graph_data, n=10)

ggplot(bar_graph_data, aes(x=TotalCost, y=InvoiceNo, fill=InvoiceNo)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top 10 least expensive orders") +
  xlab("Total Cost")
```





